The FAIRification of research in real-world evidence: A practical introduction to reproducible analytic workflows using Git and R

### 2023-11-22 manuscript/manuscript.qmd
added section on security considerations as per R#2 comment

To do so, users will need to log into their remote repository account and navigate to the section `New repository` (**Supplementary Figure 1**). This will take a user to a form to populate basic information of the new project such as the repository name (`rwe_study`) and choose basic settings such as if the repository should be private or public and if `README.md` (see @sec-repository-structure) and `.gitignore` (see @sec-stage-and-commit-changes) files should be automatically created (**Supplementary Figure 2**).

@@

If any file(s) in the study repository should not be tracked and hence not be synchronized with the remote repository (e.g., because they include confidential information that should not be visible to others like database credentials), these can be specified in the `.gitignore` file which is typically located in the root directory of the repository. To create a `.gitignore` file, a user can either auto-generate it upon initialization of the repository on GitHub (**Supplementary Figure 2**) or manually add a text file called `.gitignore` to the root directory.

### Tags (`tag`) {#tags}

@@

## Tracking changes {#tracking-changes}

@@

@@

Moreover, when working with remote repositories, it is of utmost importance to ensure that sensitive information (such as passwords or patient-level data) are not accidentally shared. To prevent this, we recommend specifying such files in the `.gitignore` file before making the very first commit or, more preferably, not storing them in the project directory in the first place.

@@

Although data sharing is unfortunately often prohibited due to privacy and legal reasons, many databases have been mapped to common data models like OMOP[@ohdsi2019book], Sentinel[@brown2020], PCORnet[@fleurence2014] and ConcePTION [@thurin2021], which makes shared code interoperable even without direct access to the same data source. In order to achieve full replicability, authors may alternatively include a sample of dummy data to enable readers to comprehend and run the analytic code. Moreover, many larger research data sources have already provided synthetic data using identical formats and variable names while protecting patient privacy (e.g., Medicare Claims Synthetic Public Use Files).\[@gonzal

>>>>

To do so, users will need to log into their remote repository account and navigate to the section `New repository` (**Supplementary Figure 1**). This will take a user to a form to populate basic information of the new project such as the repository name (`rwe_study`) and choose basic settings such as if the repository should be private or public and if `README.md` (see @sec-repository-structure) and `.gitignore` (see @sec-stage-and-commit-changes and @sec-security) files should be automatically created (**Supplementary Figure 2**).

@@

If any file(s) in the study repository should not be tracked and hence not be synchronized with the remote repository (e.g., because they include confidential information that should not be visible to others like database credentials), these can be specified in the `.gitignore` file which is typically located in the root directory of the repository. To create a `.gitignore` file, a user can either auto-generate it upon initialization of the repository on GitHub (**Supplementary Figure 2**) or manually add a text file called `.gitignore` to the root directory (more information in @sec-security).

#### Tags (`tag`) {#sec-tags}

@@

## Tracking changes {#sec-tracking-changes}

@@

## Security considerations {#sec-security}

When working with Git and remote repositories, several security considerations are crucial to safeguard sensitive information. First, it is of utmost importance to ensure that sensitive information (such as personal access tokens, passwords or patient-level data) are not accidentally shared. One way to prevent this is to configure a `.gitignore` file, which is a flat text file typically stored in the root directory of the repository. Each line in `.gitignore` specifies a file name, sub-directory name or a file name pattern (e.g., "\*.csv" for all files of type csv) that a user intentionally does not want to track and synchronize with a remote repository. If it can be avoided, such sensitive information should alternatively not be stored in the project directory in the first place.

```{bash}
# Example .gitgnore file specyfing which 
# files, sub-directories and patterns
# should not be tracked

# password .txt file
pwd.txt

# Session Data files
.RData
.Rhistory

# any type of .csv file
*.csv

# output directory
output/

# .Renviron in root directory
.Renviron
```

Second, it is good practice to not include passwords or personal access tokens inside an analytic code file but to call these from an external file. RStudio, for example, offers the possibility to create environment files (`.Renviron` and `.RProfile`) which are user-controllable text files which can be stored in the project root directory and is automatically called upon an R session start. Such a file can be configured to not be tracked using `.gitignore` (as illustrated above) and accessed as show in the following.

```{r}
# Example: configuration of database access credentials inside an .Renviron file
db_username = "JoshSmith"
db_password = "User1234"
```

These credentials can now be called inside an R script, for example, to access a database (example taken from `genieBPC` R package [@lavery2022]).

```{r}
# Example how the credentials defined in .Renviron can be used
# inside an RScript
set_synapse_credentials(
  username = Sys.getenv("db_username"), 
  password = Sys.getenv("db_password")
  )
```

Third, it needs to be ensured that access permissions to remote repositories are appropriately configured, granting only necessary access levels to collaborators unless it is intended to share the repository with the public. All major remote repositories offer the option to configure a repository to be private or public.

@@

@@

Although data sharing is unfortunately often prohibited due to privacy and legal reasons, many databases have been mapped to common data models like OMOP[@ohdsi2019book], Sentinel[@brown2020], PCORnet[@fleurence2014] and ConcePTION [@thurin2021], which makes shared code interoperable even without direct access to the same data source. In order to achieve full replicability, authors may alternatively include a sample of dummy data to enable readers to comprehend and run the analytic code. Moreover, many larger research data sources have already provided synthetic data using identical formats and variable names while protecting patient privacy (e.g., Medicare Claims Synthetic Public Use Files).[@gonzales2023]
